### 2021-02-25 yfcc100m/paper/vldb2021/design.tex
Update intro and design sections. Move appendix to eval.

In this section, we further describe VDMS design principles and implementation,
which was briefly introduced in previous work \cite{vdms-nips}.

VDMS implements a client-server architecture that handles client
requests concurrently, and coordinates query execution across
the metadata and visual data components in order to return a unified response.
Users interact with both metadata and visual data
(i.e., images, videos, feature vectors) using a unified API,
in a transactional manner.
Figure \ref{fig:arch} depicts the high-level architecture of VDMS.

The metadata component is the \textit{Persistent Memory Graph
Database} (PMGD) and the (visual) data component is our Visual Compute Module.
The Visual Compute Module enables machine-friendly enhancements to
visual data, exposing high-level abstractions to the \textit{Request Server}
for dealing with a variety of images and video formats (through OpenCV and ffmpeg),
and different methods for indexing for feature vectors
(including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
Finally, a main component of VDMS is in charge of implementing the
API and orchestrating between the PMGD and the Visual Compute Module
to serve client's requests. This component is the \textit{Request Server}.

VDMS and its components are fully available as open source projects
\footnote{https://github.com/IntelLabs/\{vdms, pmgd\}}.
We briefly describe each of the main components as follows:

@@

\subsection{Persistent Memory Graph Database}

We use the Persistent Memory Graph Database (PMGD) to provide an efficient storage
solution addressing the increasing popularity of connected data and
applications that benefit from graph-like processing.
PMGD implements an in-persistent-memory graph database, PMGD, optimized
to run on a platform equipped with persistent memory.
PMGD provides a property graph model of data storage with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
Graph represents an easier abstraction to model complex problems \cite{tao}.
Moreover, the graph model makes it very suitable for the data and
access patterns shown by visual metadata, which can be easily mapped
into application-level abstractions by developers.
For instance, abstractions like \textit{BoundingBoxes} associated to
images or videos can be easily represented using nodes and edges.
With its natural ability to extend the schema very
easily (due to the use of a property graph model),
we can support new developments in machine learning that can lead to
enhancements to existing metadata over time.
These are the reasons the team chose a graph database over a
relational database as the metadata management for the implementation of VDMS.
PMGD is designed and optimized for persistent memory technologies
like Intel Optane~\cite{IntelXPoint15}, which
promise storage providing nearly the speed of DRAM and the
durability of block-oriented storage.

The Visual Compute Module was designed and implemented to provide
an internal abstraction layer for interacting with visual data.
It enables simple visual data handling and processing (i.e., basic building block
operations like crop, resize, etc).
For traditional formats (jpg, png, tiff, mp4, etc.),
the interface is an abstraction layer over OpenCV.
However, it also provides a way to use novel formats that are
better suited for visual analytics: a novel, array-based lossless image format.
This format is built on the array data manager TileDB~\cite{TileDB} and
is well suited for images that are used in visual analytics.

VDMS provides full support for video storage and operations,
in a similar way it does for images.
This includes support for encoding, decoding, and transcoding of
\textit{mp4}, \textit{avi}, and \textit{mov} containers,
as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
This is supported through the use of either OpenCV~\cite{opencv}
or \textit{libffmpeg}\cite{ffmpeg}, or both.
All operations supported for images in VDMS are also supported at the
video and frame level of the API.
On top of that, there are a number of video-specific operations that
are supported, such as the interval operations,
enabling users to retrieve clips at different
frames-per-second (FPS) versions of the video.

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities, images, and/or videos.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as \textit{descriptors}
or \textit{visual descriptors}.
We use these terms interchangeably.
These descriptors can be classified, labeled, and used to build search indexes.
Feature Vectors support is provided through our implementation based
on high-dimensional sparse arrays, also using TileDB.
In addition, the Visual Compute Library provides a wrapper
for another high-dimensional index implementation,
Facebook's Faiss~\cite{faiss}.

\subsection{Request Server}

Developers and users of machine learning frameworks and data science
applications favor simpler interfaces to access and process data.
They cannot be expected to deal with two different ways of interacting
with information (metadata and visual data) instead of focusing on the
algorithmic parts of their pipelines.
VDMS takes care of coordinating client requests across the metadata and the
visual data, and managing multiple clients through the \textit{Request Server}.
The \textit{Request Server} is a key component of the system,
as it implements VDMS' API and coordinates request and responses from
the PMGD and Visual Compute Module subsystems.
It decomposes the user queries into
metadata and visual data requests, invokes the relevant calls behind the scene,
and returns a coherent and unified response that is easy for the user to parse
and interpret.

@@

One of the most important differentiating factor of VDMS is its interface.
VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
as first class citizens. Thus, VDMS provides an API that revolves
around visual data operations and retrieval.
VDMS API is easy to use and explicitly pre-defines certain
primitives associated with metadata, images, videos, and feature vectors.
Authors have paid particular attention to hide the complexities of our internal
implementation and up-level the API to a JSON-based API,
which is very popular across various application domains.
By defining a new JSON-based API, there is a trade-off between expressiveness
(compared to well-established query languages like SPARQL, Gremlim, or even SQL)
and the ability to natively support visual data operations.
However, we believe it is possible for our API to achieve similar levels of
expressiveness compared to more mature query languages over time.

@@

>>>>

In this section, we further describe VDMS design principles and implementation,
which was briefly introduced in previous work \cite{vdms-nips}.
VDMS implementation is fully open-sourced
\footnote{https://github.com/IntelLabs/vdms}.

@@

VDMS implements a the typical client-server architecture that handles
client queries transactionally and concurrently, as most of the most common
relational and non-relational database systems
\cite{mysql, postgresql, chang2008bigtable}.
The main difference between other data management systems and VDMS is that
it goes beyond the typical supported data types (string, integers, floats,
blobs, JSON-documents, etc), recognizing visual entities (image, videos,
feature vectors, etc) as first class citizens.
VDMS API enable users to insert, index, process, and query visual data,
as well as provides full support for inserting, indexing and querying
user-defined metadata.
Users interact with both metadata and visual data using a unified API,
in a transactional manner.

VDMS provides a property \textit{graph model} abstraction with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
This is, users interact with their objects (metadata, images, videos, etc),
as if these objects were in a connected graph.
Graph represents an easier abstraction to model complex problems,
making it very suitable for the data and access patterns shown by visual metadata,
which can be easily mapped into application-level abstractions by
developers~\cite{tao}.
For instance, abstractions like \textit{BoundingBoxes} associated to
images or videos can be easily represented using nodes and edges in a graph.
These are the reasons the team chose a graph database over a
relational database as the metadata management for the implementation of VDMS.

VDMS API provides mechanism to insert and connect Images, Videos, BoundingBoxes,
Frames, and Descriptors (feature vectors), together with any metadata
associated with the objects.
Each object is a node in the graph.
The metadata associated to each object is is modeled as
"properties" of the node in the graph.
Users can query, filter, and retrieve this objects based
on its metadata (the "properties" associated with them).
VDMS does not simply treat these objects as binary blobs of data, but rather
understands them and the type of processing that is common for them,
providing the ability to run processing on-the-fly, both at insertion and
at retrieval type.
Users can, for instance, retrieve a resized version of an image,
or retrieve only a clip from a large video.
This is one of the main differentiation aspect when compared to other
database systems, including relational databases.
In a relational database, for instance, one can query \textit{and compute}
over values in a column only for basic data types (strings, integers, floats),
and over the abstractions the relational model support (tables, columns).
A SQL query can then retrieve the \textit{computed average} "salary" of all
the employees in a company (stored using the table/columns abstraction),
but cannot perform any computation over data stored as a blob or binary object.
VDMS, because it recognized the nature of visual objects by design,
provides the ability to \textit{compute} on these visual objects
(image, videos, etc).

VDMS API also allows users to insert application-defined "Entities",
that enable applications to model any use-case specific metadata.
An "Entity" object (and its properties) is a node in the graph.
For example, a user can define an Entity object of a class "Person",
and \textit{connect} this person to one or multiple image objects.
Later, the user can retrieve the Entity object corresponding to a person,
together with all the images \textit{connected} to it.

By providing the ability to store visual data objects together with
application-defined entities and its properties, VDMS can manage
all the data the application need behind a single, unified API.
This is in contrast with current applications that rely on a combination
of multiple systems and APIs to access different portions of the data they need
\cite{tao, sculley2015hidden, mayer2020scalable, sculley2015hidden}.

Figure \ref{fig:arch} depicts the high-level architecture of VDMS, which
is composed by several subcomponents, and a Query Engine that implements
the unified API and hides all the complexity underneath.
We first describe the VDMS API, and then describe its subcomponents
and design decisions.

@@

One of the most important differentiating aspects of VDMS is its API.
VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
as first class citizens.
Thus, VDMS' API revolves around visual data operations and retrieval.
VDMS API is easy to use and explicitly pre-defines certain
primitives associated with metadata, images, videos, and feature vectors.
Authors have paid particular attention to hide the complexities of our internal
implementation and up-level the API to a JSON-based API,
which is very popular across various application domains.
By defining a new JSON-based API, there is a trade-off between expressiveness
(compared to well-established query languages like SPARQL, Gremlim, or even SQL)
and the ability to natively support visual data operations.
However, we believe it is possible for our API to achieve similar levels of
expressiveness compared to more mature query languages over time.

@@

\subsection{Graph Engine}

The Graph Engine used for in VDMS is the
\textit{Persistent Memory Graph Database} (PMGD).
PMGD was developed at Intel Labs, and was designed and optimized for
persistent memory technologies like Intel Optane~\cite{IntelXPoint15}, which
promise storage providing nearly the speed of DRAM and the
durability of block-oriented storage.
PMGD is also fully open-sourced \footnote{https://github.com/IntelLabs/pmgd}.
PMGD implements an in-persistent-memory graph database, PMGD, optimized
to run on a platform equipped with persistent memory, but
also provides the option to run on SSD and ext4 filesystem
using OS support to provide transactional guarantees (through \textit{msync}).
PMGD provides a property graph model of data storage with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
PMGD comes as a C++ library that VDMS uses for its internal query engine.
Because PMGD already supports most of the graph abstractions we wanted
to see on VDMS API, it was our preferred option.
Also, our internal evaluation shows that PMGD is faster
than other graph databases, including Neo4j\cite{miller2013graph}.
PMGD performance evaluation will be published separately.

% The Visual Compute Module enables machine-friendly enhancements to
% visual data, exposing high-level abstractions to the \textit{Request Server}
% for dealing with a variety of images and video formats (through OpenCV and ffmpeg),
% and different methods for indexing for feature vectors
% (including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
% Finally, a main component of VDMS is in charge of implementing the
% API and orchestrating between the PMGD and the Visual Compute Module
% to serve client's requests. This component is the \textit{Request Server}.

The Visual Compute Module was designed and implemented to provide
an internal abstraction layer for interacting with visual data.
It enables simple visual data handling and processing
(i.e., basic building block operations like crop, resize, etc).
For traditional formats (jpg, png, tiff, mp4, etc.),
the interface is an abstraction layer over OpenCV.
However, it also provides a way to use novel formats that are
better suited for visual analytics: a novel, array-based lossless image format.
This format is built on the array data manager TileDB~\cite{TileDB} and
is well suited for images that are used in visual analytics.
Note that, even if VDMS currently provides support for array-based lossless
image format, we do not use it as part of this evaluation.
This work focuses on a more direct comparison with a combination of alternatives,
and thus we use the traditional format (jpg and png).
The performance comparison between the array-based lossless image format
we develop and other similar formats (like PNG or TIFF) are outside of the
scope of this evaluation.

VDMS also provides full support for video storage and operations,
in a similar way it does for images.
This includes support for encoding, decoding, and transcoding of
\textit{mp4}, \textit{avi}, and \textit{mov} containers,
as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
This is supported through the use of either OpenCV~\cite{opencv}
or \textit{libffmpeg}\cite{ffmpeg}, or both.
All operations supported for images in VDMS are also supported at the
video and frame level of the API.
On top of that, there are a number of video-specific operations that
are supported, such as the interval operations,
enabling users to retrieve clips at different
frames-per-second (FPS) versions of the video.

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities, images, and/or videos.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as \textit{descriptors}
or \textit{visual descriptors}.
We use these terms interchangeably.
These descriptors can be classified, labeled, and used to build search indexes.
Feature Vectors support is provided through our implementation based
on high-dimensional sparse arrays, also using array-based approaches.
In addition, the Visual Compute Library provides a wrapper
for another high-dimensional index implementation,
Facebook's Faiss~\cite{faiss}.

### 2021-02-25 yfcc100m/paper/vldb2021/eval.tex
Update intro and design sections. Move appendix to eval.

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/all_build_sizes_plot}
\caption{Data footprint (in GB) of MySQL, PostgreSQL, and VDMS databases.}
\label{fig:db_sizes}
\end{figure}

@@

>>>>

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\columnwidth]{figures/all_build_sizes_plot}
\caption{Data footprint (in GB) of MySQL, PostgreSQL, and VDMS databases.}
\label{fig:db_sizes}
\end{figure}

@@

Figure~\ref{fig:q_throughput_16} shows the measured throughput,
as images per second, delivered by each system for all queries and databases.
In comparison to figure~\ref{fig:q_throughput_56},
when using fewer concurrent clients (nthr = 16),
the throughput for VDMS for queries \textit{q4} (\textit{2tag\_or\_resize}),
\textit{q5} (\textit{2tag\_resize\_geo\_and}), and
\textit{q6} (\textit{2tag\_resize\_geo\_or}) improved by a large margin
for the 50M and 100M databases.
In the case of \textit{q4} and \textit{q6},
the throughput for 50M (and 100M for \textit{q4})
increase by a large margin which outperforms both baselines while in
Figure~\ref{fig:q_throughput_56} PostgreSQL slightly performed better than VDMS.
In the case of \textit{q5}, all systems perform a little differently for databases
larger than 5M.
The gaps between the throughput of each system narrows at 10M.
At a database size of 50M, the performance of the baselines drastically drop
while the throughput of VDMS has a large improvement in comparison.

\begin{figure}[ht!]
\includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
\caption{Throughput Analysis using all queries from our use-case
described in the Experimental Setup Section.
We show queries in different figures for readability reasons.
The experiments show the performance of all systems (VDMS and the two baselines) as the
database size increases.
These queries were run using 16 simultaneous clients (nthr = 16),
and averaged out of 5 runs.}
\label{fig:q_throughput_16}
\end{figure}

### 2021-02-25 yfcc100m/paper/vldb2021/intro.tex
Expand on API on design section
Update intro and design sections. Move appendix to eval.

Visual computing workloads performing analytics on videos and/or images
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning (ML) techniques to
extract information from the visual data which can subsequently be used
for informed decision making \cite{vdms-nips}.
The insights this information can provide depend on the
application: retail vendors might be interested in knowing
which are the most visited areas of their stores using security
video feeds as input, or a doctor might want know the effect of a
specific treatment by looking at the changes in size of a tumor from a brain scan.

@@

We show how VDMS can be used as a centralize point for data
management and data access even when having multiple modalities of data:
Metadata, Image, Videos, and Feature Vectors.

\subsection{Related Work}

There are several databases and data management systems that focus one enabling
analytics or combining transactional and analytic workloads over large
scale data, such as SciDB~\cite{brown2010overview},
BigTable~\cite{chang2008bigtable}, Shark~\cite{xin2013shark},
and Vertica~\cite{lamb2012vertica}.
While these systems do not focus on visual processing as a primary entity,
they are valuable resources in distributing and analyzing very large scale data.
A more concrete application and case of study include
the Facebook architecture for photos that combines Tao~\cite{tao},
Haystack~\cite{beaver2010finding}, and f4~\cite{muralidhar2014f4} for metadata,
hot/recent data, and warm data respectively.
While the social aspect of Facebook is logically suited to a graph,
the bulk of their data is already stored in MySQL~\cite{mysql},
so they chose to develop a graph-aware cache.
A similar approach was taken by Grail~\cite{fan2015case},
which argued that a syntactic layer written on top of a relational database
could answer graph queries.
Diamond~\cite{huston2004diamond,satyanarayanan2010unique} exploited active
disks to perform discard-based searching order to shrink the amount of data
returned to a reasonable size.
We accomplish the same goal by exploiting persistent memory to store a
graph database for metadata.

>>>>

Visual computing workloads performing analytics on images and/or videos
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning (ML) techniques
to extract information from the visual data which can subsequently be used
for informed decision making \cite{vdms-nips}.
The insights this information can provide depend on the
application: retail vendors might be interested in knowing which are the most
visited areas of their stores using security video feeds as input,
or a doctor might want know the effect of a specific treatment by
looking at the changes in size of a tumor from a brain scan.

@@

Overall, this paper makes the following contributions:

\begin{itemize}
\item We present the design and implementation of the Visual Data Management System.
To the best of our knowledge, there are no other system capable of
managing visual data and its metadata behind a unified API
and that, at the same time, enable applications to operate over visual data
(images, videos, feature vectors).

\item We introduce the VDMS API in detail, designed for Machine Learning and visual
data analytics at scale.

\item We use our internal use-cases and implementations of visual data processing
pipeline to make a comprehensive evaluation between common ad-hoc solutions
that rely on multiple systems and VDMS.

\item We use the largest public available image dataset, YFCC100m, together
with machine generated tags for our real-world use case, providing a performance
evaluation both from the point of view of high concurrency as well as
scale (the database size, including images and metadata is 12TB in size).
\end{itemize}

\section{Related Work}

There are several databases and data management systems that focus one enabling
analytics or combining transactional and analytic workloads over large
scale data, such as SciDB~\cite{brown2010overview},
BigTable~\cite{chang2008bigtable}, Shark~\cite{xin2013shark},
and Vertica~\cite{lamb2012vertica}.
While these systems do not focus on visual processing as a primary entity,
they are valuable resources in distributing and analyzing very large scale data.
A more concrete application involving visual data include
the Facebook architecture for photos that combines Tao~\cite{tao},
Haystack~\cite{beaver2010finding}, and f4~\cite{muralidhar2014f4} for metadata,
hot/recent data, and warm data respectively.
While the social aspect of Facebook is logically suited to a graph,
the bulk of their data is already stored in MySQL~\cite{mysql},
so they chose to develop a graph-aware cache.
A similar approach was taken by Grail~\cite{fan2015case},
which argued that a syntactic layer written on top of a relational database
could answer graph queries.
Diamond~\cite{huston2004diamond,satyanarayanan2010unique} exploited active
disks to perform discard-based searching order to shrink the amount of data
returned to a reasonable size.
We accomplish the same goal by exploiting persistent memory to store a
graph database for metadata.
